---
title: "Exam 1 - Viktor"
author: "Viktor Damm"
date: "2024-05-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sandwich)
library(foreign)
library(texreg)
```

```{r}
data1 <- read.csv("~/Desktop/Oecon/4. semester/Økonometri/Econ Exam/data1.csv")
salary<-data1$salary
educ<-data1$educ
salbegin<-data1$salbegin
male<-data1$male
minority<-data1$minority
```

log(salary) = β0 + β1educ + β2log(salbegin) + β3male + β4minority + u

#1. Estimer modellen vha. OLS. Kommenter på outputtet og fortolk resultaterne.
```{r}
reg=lm(log(salary)~educ+log(salbegin)+male+minority)
summary(reg)
```
We can see that salary increases the most when the beginning salary increases with 1% holding everything else fixed.
We can also see that a male gets 4% higher salary than females holding everything else fixed.
- All are significant
- Adjusted R^square of 80%
- They are jointly significant
- All have the expected signs

#Opgave 2. - Udfør grafisk modelkontrol

We make 4 different graphic model control: 

```{r}
plot(reg, 1)
```
-The dots are equally sread around a horizontal line, indicating, that we dont have any non-linear relationship 
- This is a sign of well specified model

```{r}
plot(reg, 2)
```
- This plot shows, if the residuals are normally distributed
- If they fit nicely around the dottet line, then they are normally distributed 
- Verifies MLR 6
- You could also plot a histogram of the residuals!

```{r}
plot(reg, 3)
```
- This one checks for homoskedasticity
- -The dots are equally sread around a horizontal line, indicating, that we have homoskedasticity
- They are randomly spread and does not increase with the fitted values 

```{r}
plot(reg, 5)
```
- THis one check, if we have any outliers of major impact
_ Cook's distance lines are not even present, meaning that there are no outliers of impact. 

#3. Test for heteroskedasticitet vha. Breusch-Pagan-testet og specialudgaven af White-testet.

```{r}
residuals <- resid(reg)
residuals_squared <- residuals^2
```

```{r}
res_mod <- lm(residuals_squared ~ educ + log(salbegin) + male + minority)
summary(res_mod)
```

```{r}
multiple_r_squared <- 0.02923
n <- nrow(data1)
lm_test <- multiple_r_squared*n
lm_test
```

```{r}
p_value_lmtest <- 1-pchisq(lm_test,4)
p_value_lmtest
```

```{r}
library(lmtest)
bptest(reg)
```
- We reject H0, so there is Heteroskedasticity
- We cannot use the graphical model control

We now use the White-Test: 

```{r}
y_hat <- reg$fitted.values
y_hatsquared <- reg$fitted.values^2
white_test <- lm(residuals_squared ~ y_hat + y_hatsquared)
summary(white_test)
```
- We reject H0 again!

#4. Beregn robuste standardfejl for modellen og sammenlign med resultaterne i spørgsmål 1.

- Vi bruger R til at udregne dem, men vi skal også have formlen med, så dden kan forklares: 

```{r}
reg_robust <- coeftest(reg, vcov = vcovHC(reg, type = "HC0"))
screenreg(list(OLS =reg, OLS_robust_se =reg_robust), digits =4)
```
- Here we can compare our standard errors with our robust standard errors
- Not so big a differene
- The results does not change
- Heterosekdasticity is not very strong in this example


#5. Test hypotesen H0 : β2 = 1 mod alternativet H1 : β2 ̸= 1.
```{r}
t_stat<-(0.82180-1)/0.0374
t_stat
c<-qt(0.05, n-1)
c
p_value<- pt(t_stat, n-1)
p_value
```
- We reject the Null Hypothesis
- Beta_2 is statistically different from 1
- We reject at a 5 pct. significance level 

#6. Test hypotesen H0 : β3 = β4 = 0.

```{r}
reg_restricted <- lm(log(salary)~educ+log(salbegin))
r2_un <- summary(reg)$r.squared
r2_r <- summary(reg_restricted)$r.squared
df <- nrow(data1)-4-1
```

```{r}
#Here we are obtaining our F-statistic
f_stat <- ((r2_un - r2_r)/2)/((1 - r2_un)/df)
f_critical <- qf(0.95, 2, df)
p_value_f <- 1-pf(f_stat, 2, df)
```
- We reject our Null hypothesis 
- Beta_3 and beta_4 are statistically different from 0
- We cxan also test it with a function in R:

```{r}
library(foreign)
library(car)
myh0 <- c("male=0", "minority=0")
linearHypothesis(reg, myh0)
```

#7. Estimer modellen vha. FGLS og kommenter på resultaterne

```{r}
logu_2 <- log(resid(reg)^2) #Obtain residuals from the original regression - We square them - And we log them
```

```{r}
varreg <- lm(logu_2 ~ educ + log(salbegin) + male + minority) #We run a regression on the log_squared_residuals
```

```{r}
w <- exp(fitted(varreg)) #Obtaining the weights
```

```{r}
fgls <- lm(log(salary)~educ+log(salbegin)+male+minority, weight=1/w)
screenreg(list(OLS = reg, FGLS = fgls), digits = 4)
```
- No differencs between the estimates, standard errors or significance of the estimates 

#8. Har FGLS estimationen taget højde for al heteroskedasticiteten?

```{r}
fgls_robust <- coeftest(fgls, vcov = vcovHC(fgls, type = "HC0"))
screenreg(list(OLS = reg, FGLS = fgls, FGLS_robust = fgls_robust), digits = 4)
```
- Here we have obtained robust standard errors for the FGLS 