---
title: "Nicolaj Økonometri projekt"
author: "Nicolaj"
date: "2024-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Exam 2:
Opgave 1

#1.Estimer de to modeller vha. OLS. Kommenter på outputtet, sammenlign og fortolk resultaterne.

```{r}
library(readr)
data2 <- read_csv("data2.csv")
```
sal ar y = β0 +β1educ +β2salbeginn +β3male +β4mi nor i t y +u (1)
og
log(salar y) = β0 +β1educ +β2 log(salbeg in)+β3male +β4minor it y +u
```{r}
salary <- data2$salary
logsalary <- data2$lsalary
educ <- data2$educ
salbegin <- data2$salbegin
lsalbegin <- data2$lsalbegin
male <- data2$male
minority <- data2$minority
jobcat <- data2$jobcat
```

```{r}
salarymodel <- lm(salary~educ+salbegin+male+minority)
logsalarymodel <- lm(logsalary~educ+lsalbegin+male+minority)
summary(salarymodel);summary(logsalarymodel)
```

Noter: 
Model 1(salarymodel)

- 1 more year of education increase the salary by 1000 dollar, holding the other variables fixed.
- A starting salary increase by a 1000 dollars a year, would increase the salary in a new job in same bank by 1600 dollars a year
- If you are a male, then your salary would be 1800 dollars higher yearly than the refferal group, which is female, holding all other variables fixed.
- If you are a minority, then your salary would be 1720 dollars lower yearly than the refferal group, which is non-minority, holding all other variables fixed.
- The coeffecient of education, salary begin, male are statisical signifikant up to the 5 % signifikant level.The minority is statistical insignificant when we have a 5% signifikant level.

The adjusted r^2 means, that the 79% of the model is explained trough the model.

Model 2

- 1 more year of education increase the salary by 2,35 percent., holding the other variables fixed.
- A starting salary increase by 1% a year, would increase the salary in a new job in same bank by 0,82% a year
- If you are a male, then your salary would be 4,5% higher yearly than the refferal group, which is female, holding all other variables fixed.
- If you are a minority, then your salary would be 4,1% lower yearly than the refferal group, which is non-minority, holding all other variables fixed.
- The coeffecient of education, salary begin, male,minority are statisical signifikant up to the 5 % signifikant level.The minority is statistical insignificant when we have a 5% signifikant level.

The adjusted r^2 means, that the 80,34% of the model is explained trough the model.


If we compare the models, minority haf become signifikant, and the adjusted r^2 has slight increased.




#2. Udfør grafisk modelkontrol af de to modeller. Hvilken model vil du foretrække?

```{r}
plot(salarymodel, main = "Salary Model"); 
plot(logsalarymodel, main = "Log Salary Model") #den tager 1,2,3,5 med
```







#3. Undersøg om de to modeller er misspecificerede vha. RESET-testet.

```{r}
summary(salarymodel)
```
```{r}
summary(logsalarymodel)
```
sal ar y = β0 +β1educ +β2salbeginn +β3male +β4mi nor i t y +u (1)
og

log(salar y) = β0 +β1educ +β2 log(salbeg in)+β3male +β4minor it y +u

```{r}
y_hat <- -6.932+0.99327*educ+1.6*salbegin+1.83*male+-1.725*minority
y_hatlog <- 0.8491+0.0235*educ+0.82*lsalbegin+0.045*male+-0.041*minority
```

Nu square vi dem og cube

```{r}
y_hat_square <- y_hat^2
y_hat_cube <- y_hat^3
y_hatlog_square <- y_hatlog^2
y_hatlog_cube <- y_hatlog^3
```
Det indsætter vi så i vores model


```{r}
reg <- lm(salary~educ+salbegin+male+minority+y_hat_square+y_hat_cube)
reglog <- lm(logsalary~educ+lsalbegin+male+minority+y_hatlog_square+y_hatlog_cube)
```

```{r}
summary(reg)
```
```{r}
summary(reglog)
```

Vi kan LAVE EN RESET TEST FOR AT SE NONLINEAR

```{r}
library(lmtest)
resettest(salarymodel); resettest(logsalarymodel)
```

VI kan lave en WALDTEST

```{r}
library(lmtest)
waldtest(reg, vcov = vcov(reg), terms = c("y_hat_square", "y_hat_cube")) #tester manuelt om de hver især er = 0
waldtest(reglog, vcov = vcov(reglog), terms = c("y_hatlog_square", "y_hatlog_cube")) #tester manuelt om de hver især er = 0
```










#4. Forklar hvorfor det kunne være relevant at medtage educ2 som forklarende variabel i de to modeller. Estimer de to modeller igen hvor educ2 inkluderes (med tilhørende koefficient β5), kommenter kort på outputtet og udfør RESET-testet igen.

```{r}
salarymodel_square <- lm(salary~educ+I(educ^2)+salbegin+male+minority)
logsalarymodel_square <- lm(logsalary~educ+I(educ^2)+lsalbegin+male+minority)
screenreg(list("Salary"=salarymodel_square, "Log Salary"=logsalarymodel_square),digits=4)
```

If the education is squared it might affekt the dependent variable diffrent than the normal education varible. Education has a negativ affect on salary, however if we square the education we get that it now has a positiv affect on salary.
In Log Salary, there is little to no impact.




```{r}
library(lmtest)
resettest(salarymodel_square); resettest(logsalarymodel_square)
```
```{r}
library(lmtest)
resettest(salarymodel); resettest(logsalarymodel)
```

We still does not reject, however if we did reject the first resettest without educ2 in our model, by adding educ2 we can see our p-value increase and thus in case we orginally rejected we would not by adding educ2.



#5. Test hypotesen H0 : β1 = β5 = 0 i begge modeller (fra spørgsmål 4)


We calculate this by a function in R as well.

```{r}
library(foreign)
library(car)
myh0 <- c("educ=0", "I(educ^2)=0")
linearHypothesis(salarymodel_square, myh0)
linearHypothesis(logsalarymodel_square, myh0)
```

We can see in both models that they are jointly signifikant, however they are not sigfikant by themselves in our Log salary model.


#6. Kunne der være problemer med målefejl i de to modeller? I hvilke tilfælde vil det udgøre et problem?





```{r}
View(data1)
```

